---
title: "R Notebook"
author: "Andrea Clausen Lind"
date: "2023-12-29"
output: html_notebook
---


## Load libraries
```{r, echo=FALSE}
#library(flowCore)
library(tidyverse)
library(viridis)
#library(hrbrthemes)
library(hexbin)
library(ggridges)
#library(patchwork)
library(ggtext)
library(ggpubr)
library(scales)
library(minpack.lm)
filter <- dplyr::filter
select <- dplyr::select
```


## Load data
```{r, }
list.files("./")
path_csv = "./2024-03-08/2024-03-08_at_04-08-58pm/"
name_sub = "2024-03-08_at_04-08-58pm."
path_meta = "./Metadata_p1_240308.csv"
  
csv_plate1 <-
  list.files(path_csv, full.names = T) %>% 
  lapply(., read.delim, sep = ",") 

#summary(csv_plate1)

metadata1 <- read.csv(path_meta, sep = ",") 

```

## Format metadata
```{r}

metadata_p1 = 
  tibble(file_id = list.files(path_csv)) %>% 
  mutate(sample_id = gsub(".CSV","",file_id)) %>% 
  mutate(well = gsub(name_sub,"",sample_id))
  
metadata_p1 <- 
  merge(metadata_p1, metadata1, by.x = "well", by.y = "well") %>% 
  select(-sample_id, -file_id, -well_id)  
  
head(metadata_p1)


```

## Convert the csv data into a tidy tibble
```{r}
# initialize empty data frame
df_flow_p1 = data.frame()

# Get no. of files
no_files = length(csv_plate1)

# loop through all files and add each file data to empty data frame
for (i in 1:no_files) {
  
  # Extract file name
  well = metadata_p1$well[i]
  
  # Convert data to tibble and add file name
  file_data =
    csv_plate1[[i]] %>% 
    as_tibble() %>% 
    mutate(WELL = well)
  
  # Join with empty data frame
  df_flow_p1 = rbind(df_flow_p1,file_data)
}

head(df_flow_p1)
head(metadata_p1)

```


## Join with metadata, add row signifying plate, and join datasets
```{r}

df_full = 
  df_flow_p1 %>% 
  full_join(metadata_p1,by = c("WELL" = "well")) %>% 
  select(-X, -P01.yeast, -strain, -WELL) 


head(df_full)
tail(df_full)
```

##Plot cell size and granularity to determine cutoffs
```{r}

p_density_FSC =
  df_full %>% 
  mutate(az_cmp58 = paste0(AZ1729_uM, "_", cmp58_uM)) %>%
  ggplot(aes(FSC.HLog, color = az_cmp58, fill = az_cmp58)) +
  geom_density(alpha = 0.1, adjust = 0.0001, linewidth=0.4) +
  facet_wrap(vars(az_cmp58),nrow = 4) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  theme_bw() +
  xlab("FSC (a.u., log-scale)") +
  ggtitle("FSC.HLog. Depending on AZ1728[uM]_cmp58[uM]") +
  theme(aspect.ratio = 0.5)

p_density_FSC

p_density_SSC =
  df_full %>% 
  mutate(az_cmp58 = paste0(AZ1729_uM, "_", cmp58_uM)) %>%
  ggplot(aes(SSC.HLog, color = az_cmp58, fill = az_cmp58)) +
  geom_density(alpha = 0.1, adjust = 0.0001, linewidth=0.4) +
  facet_wrap(vars(az_cmp58),nrow = 4) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  theme_bw() +
  xlab("SSC (a.u., log-scale)") +
  ggtitle("SSC.HLog depending on AZ1728[uM]_cmp58[uM]") +
  theme(aspect.ratio = 0.5)

p_density_SSC

p_hexbin =
  df_full %>% 
  mutate(az_cmp58 = paste0(AZ1729_uM, "_", cmp58_uM)) %>%
  ggplot(aes(x = FSC.HLog, y = SSC.HLog)) +
  stat_bin_hex(bins = 90) +
  facet_wrap(vars(az_cmp58),nrow = 4) +
  scale_color_viridis(discrete = F,direction = -1, option = "A",begin = 0.2, end=0.7) +
  scale_fill_viridis(discrete = F,direction = -1, option = "A",begin = 0.2, end=0.7) +
  theme_bw() +
  xlab("FSC (a.u., log-scale)") +
  ylab("SSC (a.u., log-scale)") +
  ggtitle("FSC vs SSC depending on AZ1728[uM]_cmp58[uM]. Raw data") +
  theme(aspect.ratio = 0.5) 

p_hexbin


```
##Save plots
```{r,eval = F, fig.width=12}
ggsave(p_density_FSC, filename = "Figures/DensityFSC_raw.pdf",width = 15)
ggsave(p_density_SSC, filename = "Figures/DensitySSC_raw.pdf",width = 15)
ggsave(p_hexbin, filename = "Figures/HexbinFSC-SSC_raw.pdf",width = 15)


```


##Filter out cells below FSC and SSC threshold 
```{r}
df_full =
  df_full %>%
  filter(GRN.B.HLog > 0) %>%
  filter(RED.R.HLog > 0.5) %>%
  filter(FSC.HLog > 2.8) %>% 
  filter(SSC.HLog < 4.8)

head(df_full)
```

##Calculate mean and standard deviation
```{r}
# Calculate mean and standard deviations of all measurment points, grouping replicates
df_mean <- 
  df_full %>% 
  group_by(AZ1729_uM, cmp58_uM) %>% 
  mutate(mean_GFP = mean(GRN.B.HLog),
         sd_GFP = sd(GRN.B.HLog)) %>% 
  ungroup()

# Calculate mean and standard deviations of all measurement points, per replicate
df_mean_rep <- 
  df_mean %>% 
  group_by(AZ1729_uM, cmp58_uM, replicate) %>% 
  mutate(mean_GFP_rep = mean(GRN.B.HLog),
         sd_GFP_rep = sd(GRN.B.HLog)) %>% 
  ungroup() %>% 
  select(AZ1729_uM, cmp58_uM, replicate, mean_GFP_rep, sd_GFP_rep) %>% 
  distinct()

# Calculate mean and standard deviations of means for each replicate
df_mean_rep_mean <- 
  df_mean_rep %>% 
  group_by(AZ1729_uM, cmp58_uM) %>% 
  mutate(mean_GFP_mean = mean(mean_GFP_rep),
         sd_GFP_mean = sd(mean_GFP_rep)) %>% 
  ungroup() 


df_mean_rep_mean_full <- 
  df_mean %>% 
  group_by(AZ1729_uM, cmp58_uM) %>% 
  mutate(mean_GFP = mean(GRN.B.HLog),
         sd_GFP = sd(GRN.B.HLog)) %>% 
  mutate(mean_GFP_rep = mean(GRN.B.HLog),
         sd_GFP_rep = sd(GRN.B.HLog)) %>% 
  mutate(mean_GFP_mean = mean(mean_GFP_rep),
         sd_GFP_mean = sd(mean_GFP_rep)) %>% 
  ungroup() %>% 
  select(AZ1729_uM, cmp58_uM, replicate, mean_GFP_rep, sd_GFP_rep, GRN.B.HLog, mean_GFP_mean, sd_GFP_mean) %>%
  distinct()

head(df_full)
head(df_mean_rep_mean)
```

## Normalize Data 
Fold-change over baseline mean for each replicate
```{r}

# Calculate fold change (FC) over baseline for all measurement point per replicate 
df_mean_rep_FC <- 
  df_mean_rep %>% 
  mutate(label = paste0(cmp58_uM, "_", AZ1729_uM)) %>% 
  group_by(replicate) %>% 
  mutate(mean_GFP_rep_FC = mean_GFP_rep - mean_GFP_rep[label == "0_0"]) %>% 
  mutate(sd_GFP_rep_FC = sd(mean_GFP_rep_FC)) %>% 
  ungroup() 

# Calculate mean and standard deviations of means for each replicate
df_mean_rep_mean_FC <- 
  df_mean_rep_FC %>% 
  group_by(cmp58_uM, AZ1729_uM) %>% 
  mutate(mean_GFP_mean_FC = mean(mean_GFP_rep_FC),
         sd_GFP_mean_FC = sd(mean_GFP_rep_FC)) %>% 
  ungroup() 

head(df_mean_rep_mean_FC)

# Test plotting data. Note, adding 0.001 to x-axis values to enable log-scale.
p_linedot_cmp58_FC =
  df_mean_rep_mean_FC %>%  
  filter(AZ1729_uM == 0) %>%
  select(-AZ1729_uM) %>% 
  mutate(color="2") %>% 
  ggplot(aes(
    x = cmp58_uM+0.001,
    y = mean_GFP_mean_FC,
    ymin = mean_GFP_mean_FC - sd_GFP_mean_FC,
    ymax = mean_GFP_mean_FC + sd_GFP_mean_FC,
    color = color,
    fill = color
  )) +
  geom_line(linewidth = 1) +
  #geom_linerange() +
  geom_point(aes(y=mean_GFP_rep_FC), alpha=0.5) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1, axis.text.x = element_markdown()) +
  coord_fixed() +
  scale_x_continuous(trans = log10_trans(),
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x))) +
  #ylim(0.5,2.75) +
  #xlim(0, 0.25) +
  ylab("Fold-change GFP (a.u., log-scale)") +
  xlab("Cmp58 [uM] (log-scale)") + 
  ggtitle(paste0("GFP activation curve cmp58"))

p_linedot_cmp58_FC

# Try saving to compare to non-normalized data
ggsave(p_linedot_cmp58_FC, filename = "Figures/LineDot_cmp58_FC.pdf",width = 5)
```


#Make figures for publication.
From this data set we will make regression curves for cmp58+AZ1729

```{r}
# Make regression for Cmp58 and AZ1729 from fold change data. 
# Use 3-parametric sigmoid curves. 
# Cmp58 REGRESSION
  df_cmp58_FC <- 
  df_mean_rep_mean_FC %>% 
  filter(AZ1729_uM == 0) %>% 
  mutate(concentration = log10(cmp58_uM*10^-6+10^-9)) %>% 
  mutate(inducer = "cmp58") 

# Run regression with nlsLM, fitting data to a 4-parameter sigmodial curve 
levm.fit_cmp58 <- nlsLM(
  mean_GFP_rep_FC ~
  Amin + ((Amax - Amin) / (1 + (10^(LogEC50 - concentration))*1)),
  data = df_cmp58_FC,
  start = list(Amin = 0, Amax = 2, LogEC50 = -3))

# Show summary of the fit, and save model
summary(levm.fit_cmp58)

capture.output(summary(levm.fit_cmp58), file = "RegressionModels/RegSum_cmp58.txt")
saveRDS(levm.fit_cmp58, file = "RegressionModels/Regress_cmp58.rda")
#readRDS("RegressionModels/Regress_Acetate.rda")

# Add predicted values to data frame 
df_cmp58_FC$predict <- predict(levm.fit_cmp58)

# Make full regression curve
concrange_cmp58 <- seq(min(df_cmp58_FC$concentration), max(df_cmp58_FC$concentration), by = 0.001)
  
reg_cmp58 <- 
  data_frame(concentration = concrange_cmp58) %>% 
  mutate(predict = predict(levm.fit_cmp58, tibble(concentration = concrange_cmp58))) %>% 
  mutate(inducer = "cmp58")

head(reg_cmp58)  

#Export regression curve
write.csv(reg_cmp58,"RegressionModels/RegressCurve_cmp58.csv")


##################
# MAKE PROPRIONATE REGRESSION FOR EACH AZ1729 CONCENTRATION
df_cmp58_az_FC <- 
  df_mean_rep_mean_FC %>% 
  mutate(concentration = log10(cmp58_uM*10^-6+10^-9)) %>%  
  mutate(inducer = "cmp58") %>% 
  mutate(allosteric_mod = paste0(AZ1729_uM, " uM AZ1729")) 

head(df_cmp58_az_FC)

# Extract the AZ1729 concentrations used
az_conc <- df_cmp58_az_FC %>% select(allosteric_mod) %>% distinct() %>% pull()

# Create empty data frame for storage
df_reg_raw_full <- tibble()
df_reg_pred_full <- tibble()

# Loop across AZ1729 concentrations
for (conc in az_conc) {
  df_subset <- 
    df_cmp58_az_FC %>% 
    filter(allosteric_mod == conc)
  
  levm.fit_conc <- nlsLM(
    mean_GFP_rep_FC ~
    Amin + ((Amax - Amin) / (1 + (10^(LogEC50 - concentration))*1)),
    data = df_subset,
    start = list(Amin = 0, Amax = 2, LogEC50 = -5))

  df_subset$predict <- predict(levm.fit_conc)
  
  summary(levm.fit_conc)
  
  df_reg_raw_full <- bind_rows(df_reg_raw_full,df_subset)
  
  # Make full regression curve
  conc_range <- seq(min(df_subset$concentration), max(df_subset$concentration), by = 0.001) 
  
  df_reg_pred <- 
  data_frame(concentration = conc_range) %>% 
  mutate(predict = predict(levm.fit_conc, tibble(concentration = conc_range))) %>% 
  mutate(inducer = "cmp58") %>% 
  mutate(allosteric_mod = conc) 
  
  df_reg_pred_full <- bind_rows(df_reg_pred_full,df_reg_pred)
  
  #Save summary and model
  capture.output(summary(levm.fit_conc), file = paste0("RegressionModels/RegSum_cmp58_",conc,".txt"))
  saveRDS(levm.fit_conc, file = paste0("RegressionModels/Regress_cmp58_",conc,".rda"))
  write.csv(df_reg_pred,paste0("RegressionModels/RegressCurve_cmp58_",conc,".csv"))

}

# Save data frame with all predicted regression curves and the raw data respectively combined
write.csv(df_reg_pred_full, paste0("RegressionModels/RegressCurve_cmp58_AZ1729_full.csv"))
write.csv(df_reg_raw_full, paste0("RegressionModels/RawData_cmp58_AZ1729_full.csv"))


# PLOT REGRESSION

# Plot cmp58 alone
p_linedot_FC_reg_cmp58 <- 
  df_cmp58_FC %>%  
  ggplot(aes(
    x = concentration,
    color = inducer,
    fill = inducer
  )) +
  geom_line(data = reg_cmp58, aes(y = predict), linewidth = 1) +
  geom_linerange(aes(y = mean_GFP_mean_FC,
                     ymin = mean_GFP_mean_FC - sd_GFP_mean_FC,
                     ymax = mean_GFP_mean_FC + sd_GFP_mean_FC)) +
  geom_point(aes(y = mean_GFP_mean_FC), size = 1.7,shape = 21, fill = "white") +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 2, axis.text.x = element_markdown()) +
  coord_fixed() +
  ylim(-0.1,1.25) +
  ylab("GFP over baseline (Log−scale))") +
  xlab("Log10([Cmp58]) M") + 
  ggtitle(paste0("GFP activation curve cmp58 ")) 

p_linedot_FC_reg_cmp58


#Plot cmp58 with different AZ1729 concentrations
p_linedot_FC_reg_cmp58_az_wide =
  df_reg_raw_full %>%  
  ggplot(aes(
    x = concentration,
    color = allosteric_mod,
    fill = allosteric_mod
  )) +
  geom_line(data = df_reg_pred_full, aes(y = predict), linewidth = 1) +
  geom_linerange(aes(y = mean_GFP_mean_FC,
                     ymin = mean_GFP_mean_FC - sd_GFP_mean_FC,
                     ymax = mean_GFP_mean_FC + sd_GFP_mean_FC,
                     group = label)) +
  geom_point(aes(y = mean_GFP_mean_FC,
                 group = label), 
             size = 1.7,
             shape = 21, 
             fill = "white") +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1, axis.text.x = element_markdown()) +
  coord_fixed() +
  ylim(-0.1,1.25) +
  ylab("GFP over baseline (Log−scale))") +
  xlab("Log10([Cmp58]) M") + 
  ggtitle(paste0("GFP activation curve cmp58 at different AZ1729 conc.")) + 
  labs(color = "AZ1729 uM", fill = "AZ1729 uM")

p_linedot_FC_reg_cmp58_az_wide

p_linedot_FC_reg_cmp58_az =
  df_reg_raw_full %>%  
  ggplot(aes(
    x = concentration,
    color = allosteric_mod,
    fill = allosteric_mod
  )) +
  geom_line(data = df_reg_pred_full, aes(y = predict), linewidth = 1) +
  geom_linerange(aes(y = mean_GFP_mean_FC,
                     ymin = mean_GFP_mean_FC - sd_GFP_mean_FC,
                     ymax = mean_GFP_mean_FC + sd_GFP_mean_FC,
                     group = label)) +
  geom_point(aes(y = mean_GFP_mean_FC,
                 group = label), 
             size = 1.7,
             shape = 21, 
             fill = "white") +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.2, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 2, axis.text.x = element_markdown()) +
  coord_fixed() +
  ylim(-0.1,1.25) +
  ylab("GFP over baseline (Log−scale))") +
  xlab("Log10([Cmp58]) M") + 
  ggtitle(paste0("GFP activation curve cmp58 at different AZ1729 conc.")) + 
  labs(color = "AZ1729 uM", fill = "AZ1729 uM")

p_linedot_FC_reg_cmp58_az

```

## Save plot
```{r,eval = F, fig.width=12}

ggsave(p_linedot_FC_reg_cmp58, filename = "FiguresNew/reg_Cmp58.pdf",height = 4)
ggsave(p_linedot_FC_reg_cmp58_az, filename = "FiguresNew/reg_Cmp58_AZ1729.pdf",height = 4)
ggsave(p_linedot_FC_reg_cmp58_az_wide, filename = "FiguresNew/reg_Cmp58_AZ1729_wide.pdf",height = 4)

```


## Calculate significace and plot barplot for Cmp58
```{r}

#PLOT AZ1729 DATA
p_barpoint_FC_reg <- 
  df_cmp58_az_FC %>%  
  mutate(cmp58_uM = as.factor(cmp58_uM),
         AZ1729_uM = as.factor(as.integer(log10(AZ1729_uM*10^-6+10^-9)))) %>% 
  ggplot(aes(
    x = cmp58_uM,
    y = mean_GFP_mean_FC
  )) +
  geom_col(aes(
    color = AZ1729_uM,
    fill = AZ1729_uM), 
    width = 0.4, 
    position = position_dodge(0.5)) +
  geom_jitter(aes(y = mean_GFP_rep_FC,
                 group = AZ1729_uM),
             alpha = 0.6,
             size = 0.6,
             position = position_dodge2(0.5)) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1, 
        axis.text.x = element_markdown()) +
  coord_fixed() +  
  ylim(-0.5,1.25) +
  ylab("GFP over baseline (log-scale)") +
  xlab("Cmp58 [uM]") + 
  ggtitle(paste0("GFP activation curve propionate"))+
  labs(color = "log10([AZ1729]) M", fill = "log10([AZ1729]) M")

p_barpoint_FC_reg

ggsave(p_barpoint_FC_reg, filename = "FiguresNew/barpoint_Cmp58_AZ1729.pdf",height = 4)

#CALCULATE SIGNIFICANCE AZ1729
conc_prop <-
  df_cmp58_az_FC %>%
  select(cmp58_uM) %>% 
  unique() %>% 
  as_vector()

conc_AZ1729 <-
  df_cmp58_az_FC %>%
  select(AZ1729_uM) %>% 
  filter(AZ1729_uM != 0) %>% 
  unique()  %>% 
  as_vector()

# Make empty frame for p-values
df_sig_AZ1729 <- 
  df_cmp58_az_FC %>% 
  select(cmp58_uM, AZ1729_uM) %>% 
  filter(AZ1729_uM != 0) %>% 
  distinct() %>% 
  arrange(cmp58_uM, AZ1729_uM) %>% 
  mutate(pval.ttest.oneside = rep(NA, length(conc_prop)*length(conc_AZ1729))) %>% 
  mutate(pval.ttest.twoside = rep(NA, length(conc_prop)*length(conc_AZ1729))) 
  

# loop through all concentrations of AZ1729 and run one-sided t-test
pos = 1

  for (i in 1:(length(df_sig_AZ1729$pval.ttest.oneside))) {
    df_ind <-
      df_cmp58_az_FC %>% 
      filter(cmp58_uM == df_sig_AZ1729$cmp58_uM[i]) %>% 
      filter(AZ1729_uM %in% c(0, df_sig_AZ1729$AZ1729_uM[i]))
    
    #paired t-test
    df_ttest_oneside <- 
      as_data_frame(df_ind) %>%
      summarize(tidy(t.test(
        mean_GFP_rep_FC~AZ1729_uM,
        paired=TRUE, 
        conf.level=0.99,
        alternative = "less"
        )))
    
    df_sig_AZ1729$pval.ttest.oneside[pos] <- df_ttest_oneside$p.value
    
    #paired t-test, one-sided
    df_ttest_twoside <- 
      as_data_frame(df_ind) %>%
      summarize(tidy(t.test(
        mean_GFP_rep_FC~AZ1729_uM,
        paired=TRUE, 
        conf.level=0.99
        )))
    
    df_sig_AZ1729$pval.ttest.twoside[pos] <- df_ttest_twoside$p.value
    
    pos = pos+1
  }

head(df_sig_AZ1729)

write_csv(df_sig_AZ1729, file = "FiguresNew/t-test_Cmp58_AZ1729.csv")

```
