---
title: "Biolector analysis"
author: "Andrea Clausen Lind"
date: "12/09/2023"
output: html_notebook
---

## Load libraries
```{r}
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(broom)
library(patchwork)
library(ggridges)
filter <- dplyr::filter
```

## Load data 
```{r}
df_data_filterset = 
  read.csv2("2023-06-01_FFAR2.csv", sep = ",", skip = 11) #Remove first 11 rows as these contain information we dont need. Keep the filter setting data.

df_data_raw = 
  read.csv2("2023-06-01_FFAR2.csv", sep = ",", skip = 22, header = F) #Remove first 22 rows to extract the readout data. 

df_meta =
  read.csv("2023-06-01_Metadata.csv") %>% 
  filter(strain != "")


head(df_data_raw)
head(df_meta)
```

##Begin by processing raw data
```{r}
df_data <- df_data_raw[3:nrow(df_data_raw),] #Start with processing of the measurement data


df_data <- df_data %>% 
  filter(V4 != "") %>%  #Remove empty rows in filterchannel
  filter(!grepl("Cal",V4)) #Remove rows with Cal.Biomass measurements as we dont use these

df_well_info <- df_data[2:nrow(df_data),1:4]

sample_col <- 
  df_well_info %>% 
  mutate(sample = paste0(V1,"_",V4)) %>% 
  pull(sample)

sample_col <- c("time_h",sample_col)

df_data$V4 <- sample_col

df_data <- df_data[,-c(1:3)]

df_data <- t(df_data) #Transpose data

row.names(df_data) = NULL #Remove row names

colnames(df_data) = df_data[1,] #Replace column names with row 1

df_data = df_data[-1,] #remove row 1

# tidy up into long format

tbl_data <- 
  df_data %>% 
  as_tibble() %>% 
  pivot_longer(cols = -time_h,names_to = "well_channel",values_to = "value") %>% 
  separate(col = well_channel,into = c("well","channel"),sep = "_") %>% 
  relocate(time_h, .after = channel)

rm(df_data,df_well_info) #Remove things we will not use again
gc()
```

##Extract well channel information
```{r}
df_channel <- 
  df_data_filterset %>% 
  select(FILTERSET,FILTERNAME,GAIN) %>% 
  slice_head(n=6) %>% 
  mutate(FILTERNAME = gsub(" .*","",FILTERNAME)) %>% # Removes weird tab chars
  mutate(FILTERNAME = ifelse(grepl("UV",FILTERNAME),"UVGFP",FILTERNAME)) %>% 
  mutate(channel_gain = paste(FILTERNAME,GAIN,sep = "_"))

colnames(df_channel) <- c("channel","channel_name","gain","channel_gain")
```

##Merge with data
```{r}
tbl_data <- 
  tbl_data %>% 
  full_join(df_channel,by = "channel") %>% 
  select(well,channel_gain,time_h,value)
```

##Join with metadata
```{r}
tbl_full <-
  tbl_data %>%
  full_join(df_meta, by = "well")

#change column type
tbl_full <-
tbl_full %>% 
  mutate(time_h = as.numeric(time_h)) %>% 
  mutate(value = as.numeric(value)) %>% 
    mutate(induction = factor(induction, levels = c("none", "1 nM Cmp1", "10 mM prop.", "10 mM prop. + 1 nM Cmp58", "1 uM Sc-pep")))

```


##Plot data overview
```{r, fig.width=7}
tbl_full %>% 
  mutate(group_rep = paste0(strain, "_" ,replicate,"_",induction)) %>% 
  ggplot(aes(x = time_h, y = value, color = induction, group = group_rep)) +
           geom_line() +
           facet_grid(vars(channel_gain), vars(strain))
```

##Correct for cell density 
```{r}
tbl_plot <-
  tbl_full %>% 
    pivot_wider(names_from = channel_gain, values_from = value) %>% 
    mutate(GFP_ODnorm = GFP_100/Biomass_30) 

#calculate mean and sd for each timepoint
tbl_plot_mean <- 
  tbl_plot %>% 
  group_by(induction, strain, time_h) %>% 
  mutate(GFP_ODnorm_mean = mean(GFP_ODnorm),
         GFP_ODnorm_sd = sd(GFP_ODnorm)) %>% 
  ungroup() %>% 
  select(strain, induction, time_h, GFP_ODnorm, GFP_ODnorm_mean, GFP_ODnorm_sd) %>%
  distinct()

```

##Plot overview
```{r}

  tbl_plot_mean %>% 
  filter(time_h < 15) %>% 
  mutate(group_rep2 = paste0(strain,"_",induction)) %>%   
  ggplot(aes(x = time_h, y = GFP_ODnorm_mean, group = group_rep2, color = induction)) + 
  geom_ribbon(aes(ymin = GFP_ODnorm_mean - GFP_ODnorm_sd,
                  ymax = GFP_ODnorm_mean + GFP_ODnorm_sd,
                  fill = induction),
              alpha =0.3,
              color = NA) +
  geom_line(size = 1) +
  facet_wrap(vars(strain)) +
  theme_bw() +
  theme(aspect.ratio = 0.6) +
  xlab("Time [h]") +
  ylab("GFP / OD600 (a.u.)") 

  tbl_plot_mean %>% 
  filter(time_h < 15) %>% 
  filter(strain != "ASC201") %>% 
  mutate(group_rep2 = paste0(strain,"_",induction)) %>%   
  ggplot(aes(x = time_h, y = GFP_ODnorm_mean, group = group_rep2, color = induction)) + 
  geom_ribbon(aes(ymin = GFP_ODnorm_mean - GFP_ODnorm_sd,
                  ymax = GFP_ODnorm_mean + GFP_ODnorm_sd,
                  fill = induction),
              alpha =0.3,
              color = NA) +
  geom_line(size = 1) +
  facet_wrap(vars(strain)) +
  theme_bw() +
  theme(aspect.ratio = 0.6) +
  xlab("Time [h]") +
  ylab("GFP intensity / OD600 [A.U.]") 



```

##GFP per OD peaks at 3 h, look further into this time point
```{r}
tbl_plot_mean_3h <- 
    tbl_plot_mean %>% 
  filter(time_h %in% c("3.04498"))

p_bar_cont <-
  tbl_plot_mean_3h %>% 
  mutate(group_rep2 = paste0(strain,"_",induction)) %>% 
  mutate(group_rep2 = as.factor(group_rep2)) %>% 
  ggplot(aes(x = strain, y=GFP_ODnorm_mean, fill = induction, group = induction)) +
    geom_bar(
    aes(fill = induction),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(y = GFP_ODnorm, 
        group = induction),
    size = 0.5,
    color = "black",
    position = position_dodge2(width = 0.9)
  ) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  xlab("Strain") +
  ylab(expression("GFP/OD600 [a.u.]")) +
  ggtitle("Induction at 3 h of incubation")
  
p_bar_cont

p_bar <-
  tbl_plot_mean_3h %>% 
  filter(strain != "ASC201") %>%
  mutate(group_rep2 = paste0(strain,"_",induction)) %>% 
  mutate(group_rep2 = as.factor(group_rep2)) %>% 
  ggplot(aes(x = strain, y=GFP_ODnorm_mean, fill = induction, group = induction)) +
    geom_bar(
    aes(fill = induction),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(y = GFP_ODnorm, 
        group = induction),
    size = 0.5,
    color = "black",
    position = position_dodge2(width = 0.9)
  ) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  xlab("Strain") +
  ylab(expression("GFP/OD600 [a.u.]")) +
  ggtitle("Induction at 3 h of incubation")
  
p_bar

# Plot ASC4G1-3
p_bar_G1 <-
  tbl_plot_mean_3h %>% 
  filter(strain == "ASC4G1-3") %>%
  mutate(group_rep2 = paste0(strain,"_",induction)) %>% 
  mutate(group_rep2 = as.factor(group_rep2)) %>% 
  ggplot(aes(x = induction, y=GFP_ODnorm_mean, fill = induction)) +
  geom_bar(aes(),
           stat = "summary",
           fun = "mean",
           width = 0.5) +
  geom_point(aes(y = GFP_ODnorm),
          size = 0.5,
          color = "black",
          position = position_dodge2(width = 0.5)) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust=1)) +
  xlab("Induction") +
  ylab(expression("GFP/OD600 [a.u.]")) +
  ggtitle("Induction at 3 h of incubation for chimeric GPA1(i1)")
  
p_bar_G1

```
##Save plots
```{r, eval=F}
ggsave(p_bar_cont, filename = "Figures/BarDotCont.pdf",width = 6)
ggsave(p_bar, filename = "Figures/BarDot.pdf",width = 6)
ggsave(p_bar_G1, filename = "Figures/BarDot_G1.pdf",width = 5)
```

##Calculate significance
```{r}
#Extract number of strains and induction combinations
all_strains <- 
  tbl_plot_mean_3h %>% 
  filter(strain != "ASC201") %>% 
  filter(induction != "none") %>% 
  mutate(induction = as.character(induction)) %>% 
  mutate(strain_ind = paste0(strain,"_",induction)) %>% 
  select(strain_ind, strain, induction) %>% 
  distinct() 

# Create result data frame
df_res <- tibble(strain_ind = all_strains$strain_ind,
                 pval.ttest = rep(NA,length(all_strains$strain_ind)),
                 )

# loop through all strains and run one-sided t-test
for (i in 1:length(all_strains$strain_ind)) {
  df_strain <- tbl_plot_mean_3h %>% 
    filter(strain != "ASC201") %>% 
    mutate(induction = as.character(induction)) %>% 
    filter(strain == all_strains$strain[i])
  
  df_induc <- 
    df_strain %>% 
    filter(induction %in% c("none",all_strains$induction[i])) %>% 
    mutate(induction = as.factor(induction)) %>%
    group_by(induction) 

  #paired t-test
  df_ttest <- 
    as_tibble(df_induc) %>%
    summarize(tidy(t.test(
      GFP_ODnorm~induction, 
      paired=TRUE, 
      conf.level=0.99,
      alternative = "greater")))
  
  df_res$pval.ttest[i] <- df_ttest$p.value
}



# Add comparison between Cmp58 addition/no addition 
Cmp58_strains <- 
  tbl_plot_mean_3h %>% 
  filter(strain != "ASC201") %>% 
  filter(!induction %in% c("none", "1 nM Cmp1", "10 mM prop.")) %>% 
  mutate(induction = as.character(induction)) %>% 
  mutate(strain_ind = paste0(strain,"_",induction)) %>% 
  select(strain_ind, strain, induction) %>% 
  distinct() 

df_res_Cmp58 <- tibble(strain_ind = paste0(Cmp58_strains$strain_ind,"_propComp"),
                 pval.ttest = rep(NA,length(Cmp58_strains$strain_ind)))


# loop through all strains and run one-sided t-test
# compare propionate with and without allosteric modulator
# One-sided "less" due to order of comparison being reversed
for (i in 1:length(Cmp58_strains$strain_ind)) {
  df_strain <- tbl_plot_mean_3h %>% 
  filter(!induction %in% c("none", "1 nM Cmp1")) %>% 
    mutate(induction = as.character(induction)) %>% 
    filter(strain == Cmp58_strains$strain[i])
  
  df_induc <- 
    df_strain %>% 
    filter(induction %in% c("10 mM prop.",Cmp58_strains$induction[i])) %>% 
    mutate(induction = as.factor(induction)) %>%
    group_by(induction) 

  #paired t-test
  #
  df_ttest <- 
    as_tibble(df_induc) %>% 
    summarize(tidy(t.test(
      GFP_ODnorm~induction, 
      paired=TRUE, 
      conf.level=0.99,
      alternative = "less")))
  
  df_res_Cmp58$pval.ttest[i] <- df_ttest$p.value
}

#Summarize the p-value results
df_res_final <-
  rbind(df_res, df_res_Cmp58)

#Save results as csv
write_csv(df_res_final, file = "Figures/t-test_BarDot")

# Add comparison between baselines in the uninduced state (anova)
baseline_strains <- 
  tbl_plot_mean_3h %>% 
  filter(strain != "ASC201") %>% 
  filter(induction %in% c("none")) %>% 
  mutate(induction = as.character(induction)) %>% 
  mutate(strain_ind = paste0(strain,"_",induction)) %>% 
  select(strain_ind, strain, induction) %>% 
  distinct() 

df_res_base <- tibble(strain_ind = baseline_strains$strain_ind,
                 pval.anova = rep(NA,length(baseline_strains$strain_ind)))

df_strain <- tbl_plot_mean_3h %>% 
  filter(induction %in% c("none")) %>% 
  mutate(induction = as.character(induction)) %>% 
  group_by(strain)
  
  #anova 
  df_anova <- 
    as_tibble(df_strain) %>% 
    reframe(tidy(aov(GFP_ODnorm~strain)))



#Save results as csv
write_csv(df_anova, file = "Figures/Anova_BarDot_baseline")

```


##Generate plots for publication
```{r}

tbl_plot_induction <- 
  tbl_plot_mean %>% 
  filter(time_h < 10) %>% 
  filter(strain != "ASC201") %>% 
  mutate(group_rep2 = paste0(strain,"_",induction)) %>%   
  ggplot(aes(x = time_h, y = GFP_ODnorm_mean, group = group_rep2, color = induction)) + 
  geom_ribbon(aes(ymin = GFP_ODnorm_mean - GFP_ODnorm_sd,
                  ymax = GFP_ODnorm_mean + GFP_ODnorm_sd,
                  fill = induction),
              alpha =0.3,
              color = NA) +
  geom_line(size = 1) +
  facet_wrap(vars(strain)) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 0.6) +
  xlab("Time [h]") +
  ylab("GFP / OD600 (a.u.)") 

ggsave(tbl_plot_induction, filename = "FiguresNew/timelineInducG-prot.pdf",height = 4)

p_bar <-
  tbl_plot_mean_3h %>% 
  filter(strain != "ASC201") %>%
  mutate(group_rep2 = paste0(strain,"_",induction)) %>% 
  mutate(group_rep2 = as.factor(group_rep2)) %>% 
  ggplot(aes(x = strain, y=GFP_ODnorm_mean, fill = induction, group = induction)) +
    geom_bar(
    aes(fill = induction),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(y = GFP_ODnorm, 
        group = induction),
    size = 0.5,
    color = "black",
    position = position_dodge2(width = 0.9)
  ) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  ylim(0,4) +
  xlab("Strain") +
  ylab(expression("GFP/OD600 [a.u.]")) +
  ggtitle("Induction at 3 h of incubation")
  
p_bar

ggsave(p_bar, filename = "FiguresNew/BarDot-FFA2.pdf",height = 4)
```


##Look into if there is any transient activation of GFP in the negative control strain
```{r}
tbl_plot_mean_cont <- 
  tbl_plot_mean  %>% 
  filter(induction == "none")

p_bar_cont_transient <-
  tbl_plot_mean_cont %>% 
  ggplot(aes(x = time_h, y=GFP_ODnorm_mean)) +
    geom_bar(
    aes(fill = time_h),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(y = GFP_ODnorm),
    size = 0.5,
    color = "black"
  ) +
  facet_wrap(vars(strain),nrow = 1) +
  scale_color_viridis(discrete = F,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = F,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  xlab("Time") +
  ylab(expression("GFP/OD600 [a.u.]")) +
  ggtitle("Transient induction for control contition")
  
p_bar_cont_transient




# Look into if there is significant difference between t=0 and t=3 for the control
cont_strain <- 
  tbl_plot_mean %>% 
  filter(strain != "ASC201") %>% 
  filter(induction == "none") %>% 
  #mutate(time_h = as.character(time_h)) %>% 
  mutate(strain_h = paste0(strain,"_",time_h)) %>% 
  select(strain_h, strain, time_h, GFP_ODnorm, GFP_ODnorm_mean) %>% 
  distinct() %>% 
  filter(time_h %in% c("0.04218","1.04307","2.04392","3.04498", "4.04564")) 




# Plot data to check
p_bar_cont_check <-
  cont_strain %>% 
  ggplot(aes(x = time_h, y=GFP_ODnorm)) +
    geom_bar(
    aes(fill = time_h, group = time_h),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(y = GFP_ODnorm),
    size = 0.5,
    color = "black"
  ) +
  facet_wrap(vars(strain),nrow = 1) +
  scale_color_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  scale_fill_viridis(discrete = T,direction = -1, option = "A",begin = 0.3, end=0.8) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  xlab("Time") +
  ylab(expression("GFP/OD600 [a.u.]")) +
  ggtitle("Transient induction for control contition")
  
p_bar_cont_check



  df_ttest <- 
    cont_strain %>% 
    filter(strain == "ASC4G1-3") %>% 
    filter(time_h %in% c("0.04218","3.04498")) %>% 
    as_tibble() %>%
    summarize(tidy(t.test(
      GFP_ODnorm~time_h, 
      paired=TRUE, 
      conf.level=0.99,
      alternative = "less"
      )))

```




